---
title: "Exploring the gapminder data"
author: "Kejing Li"
date: "11/24/2019"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---
# Relationship between population density and average life expectancy 
```{r setup, include=FALSE}
# Set global options
# Hide all codes, messages and warnings
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
# Set figure options for better display                      
                      fig.width = 6, fig.asp = 0.618,
                      fig.align = "center") 
# Set numerical output display parameters
options(width = 70, digits = 2, scipen = 7)

#load required packages
library(tidyverse)
library(geonames)
library(gapminder)
library(countrycode)
library(broom)
library(modelr)
library(knitr)

#set default theme and modify the font size
theme_set(theme_bw(base_size = 10))
```

```{r merge, include = FALSE, cache = TRUE}
#get information from geonames, to do this you need first have an API key(here, an valid account name) in Rprofile 
country_info <- GNcountryInfo() %>%
  as_tibble()

#unify the format of country name and the variable name to create a unique key for merging 
country_info$countryCode <- countrycode(country_info$countryCode,
                                        origin = "iso2c", 
                                        destination = "country.name")
country_info_merge <- country_info %>%
  rename( "country"= "countryCode")


#use left join to keep the observation of gapminder 
data <-  gapminder %>%
           left_join(country_info_merge, 
                     by = "country") 

#select and calculate variables for analysis
gapminder_analysis <- data%>%
  select("country","lifeExp","pop","areaInSqKm")%>%
  mutate(areaInSqKm = as.numeric(areaInSqKm),
         pop_density = pop/areaInSqKm)
#rename variables in a consistent way
colnames(gapminder_analysis) <- c("country", "life_exp","pop","area", "pop_density")
```

## 1.Graph 

```{r graph}
gapminder_analysis %>%
  ggplot(aes(pop_density,life_exp))+
  geom_point(alpha = .1)+
  geom_smooth() +
#zoom in to 0-500 because data are mostly clustered at small density values
  scale_x_continuous(limits = c(0, 500))+
  labs(x = "Population Density (per square kilometer)",
       y = "Average Life Expectancy",
       caption = "Source: Gapminder, GeoNames geographical database")
```
The graph suggests when population density is below around 100 million people/$km^2$, the average life expectancy seems to increase with the population density; when population density exceeds 100 million/$km^2$, there’s a weak relationship. 

## 2.Model 
To confirm the relationship, I fit a single variable linear regression model of the relationship between the population density and average life expectancy. The coefficient estimates are as follows:

```{r model}
# Fit a linear model for the data
model <- lm(life_exp ~ pop_density, data = gapminder_analysis)
# Find and display the model's estimates
model_coef <- tidy(model, conf.int = TRUE) %>%
# Change the labels
mutate(term = recode(term, pop_density = "Population Density"))
kable(model_coef)
```


Though the predictor has a low *p*-value, the coefficiency is only 0.01. Based on the model, population density may not be a good estimator for the average life expectancy. 