---
title: "NASA - Near Earth Asteroid Information"
author: "Kejing Li"
date: "11/24/2019"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---
# Introduction

NeoWs (Near Earth Object Web Service) is a RESTful web service for near earth Asteroid information. With NeoWs a user can: search for Asteroids based on their closest approach date to Earth, lookup a specific Asteroid with its NASA JPL small body id, as well as browse the overall data-set.[^1] In this report, I use NeoWs API to get overall data-set of asteroids. And the analysis will be around two variables:

- Magnitude of asteroids
- Is potential hazardous asteroids

As stated on the NASA websites:all asteroids with a minimum orbit intersection distance (MOID) of 0.05 au or less and an absolute magnitude (H) of 22.0 or less are considered PHAs.[^2] Here,a confirmation and extension of the statement on the relationship between the absolute magnitude and PHAs from the actual data set is demonstrated. 

```{r setup, include=FALSE}
# Set global options
# Hide all codes, messages and warnings
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
# Set figure options for better display                      
                      fig.width = 6,
                      fig.asp = 0.618,
                      fig.align = "center",
                      out.width = "70%",
                      out.height = "70%")

#load required packages
library(tidyverse)
library(httr)
library(modelr)

#set default theme and modify the font size
theme_set(theme_bw(base_size = 13))
```

```{r}
#get api key stored in the Rprofile
nasa_key <- getOption("nasa_key")
#browse the overall Asteroid data-set
nasa <- GET("https://api.nasa.gov/neo/rest/v1/neo/browse?",
            query = list(api_key = nasa_key))
#remove links and pages which are of different length 
asteroid <- content(nasa) %>%
    list_modify(links = NULL,
                page = NULL) %>%
    as.tibble()

#unnest the lists
asteroid <- 
  asteroid %>% hoist(near_earth_objects,
                     absolute_magnitude_h = "absolute_magnitude_h",
                     is_potentially_hazardous_asteroid = "is_potentially_hazardous_asteroid")
```

# Overview 
To get an overview of the near earth asteroids, I plot the distribution and number of magnitude and hazardousness separately.
 
## Magnitude 
**H (absolute magnitude)**:An asteroid’s absolute magnitude is the visual magnitude an observer would record if the asteroid were placed 1 Astronomical Unit (au) away, and 1 au from the Sun and at a zero phase angle.[^3] In short,the absolute magnitude of an astronomical object is a measure of its intrinsic light output, independent of its distance. 

```{r}
#count the magnitude
asteroid %>%
  ggplot()+
  geom_freqpoly(aes(absolute_magnitude_h))+
  labs(x = "Absolute Magnitude",
       y = "Number",
       caption = "Source:NASA JPL Asteroid team")
```

The graph suggests that the absolute magnitude of near earth asteroids is between 15-27.

## Hazardousness
**PHA (Potentially Hazardous Asteroid)**:Potentially Hazardous Asteroids (PHAs) are currently defined based on parameters that measure the asteroid’s potential to make threatening close approaches to the Earth. [^4]

```{r}
#count the hazardousness
asteroid %>%
  count(is_potentially_hazardous_asteroid)%>%
  ggplot(aes(is_potentially_hazardous_asteroid, n))+
  geom_col(width = 0.3)+
  geom_text(aes(label = n),vjust = "top", color= "white")+
  labs(x = "PHA (Potentially Hazardous Asteroid)",
       y = "Number",
       caption = "Source:NASA JPL Asteroid team")
```

The graph suggests, only 2 of the near earth asteroids are considered hazardous. 

# Relationship
To confirm the stated relationship between the magnitude and hazardousness, here, I put them into the same plot.
```{r}
asteroid %>%
  ggplot(aes(is_potentially_hazardous_asteroid, absolute_magnitude_h))+
  geom_boxplot()+
  labs(x = "Hazardousness of asteroids",
       y = "Magnitude of asteroids",
       caption = "Source:NASA JPL Asteroid team")
```

From the graph, it is true that all the hazardousness asteroids are below the absolute magnitude (H) of 22.0. But not all asteroids that are not considered hazardous are over the absolute magnitude (H) of 22.0, this is because some of the asteroids below the absolute magnitude (H) of 22.0 may not have a minimum orbit intersection distance (MOID) of 0.5 or less.

# Conclusion 
Overall, we prove that the smaller absolute magnitude of an asteroid is, the more likely it is a potentially hazardous asteroid. This relationship can be identified in the graph below:
```{r}
asteroid %>%
  mutate(is_potentially_hazardous_asteroid = as.numeric(is_potentially_hazardous_asteroid))%>% 
  ggplot(aes(absolute_magnitude_h, is_potentially_hazardous_asteroid))+
   scale_y_continuous(labels = scales::percent)+
   geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              se = FALSE)+
  labs(x = "Absolute Magnitude",
       y = "Probability of being hazardous",
       caption = "Source:NASA JPL Asteroid team")

```



[^1]: "Asteroids - NeoWs API.", Nasa open data portal, July 19, 2018. https://data.nasa.gov/Space-Science/Asteroids-NeoWs-API/73uw-d9i8.
[^2]: "Glossary - PHA (Potentially Hazardous Asteroid)", Jet Propulsion Lab
https://cneos.jpl.nasa.gov/glossary/PHA.html
[^3]: "Glossary - H (absolute magnitude)", Jet Propulsion Lab
https://cneos.jpl.nasa.gov/glossary/h.html
[^4]: "Glossary - PHA (Potentially Hazardous Asteroid)", Jet Propulsion Lab
https://cneos.jpl.nasa.gov/glossary/PHA.html

